Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-11-18 grove-paper.tex
Replace '*' with `\ast`

% Load MnSymbol without clobbering \ast
% See https://tex.stackexchange.com/a/269691
\usepackage{amsmath}% needed before mathabx
\let\amsast=\ast
\usepackage[matha]{mathabx}% needed to prevent \ast getting clobbered
\let\abxast=\ast
\usepackage{MnSymbol}
\let\mnast=\ast
\let\ast=\abxast
%\def\*{\ast}
\DeclareMathSymbol{*}{\mathbin}{symbols}{"03} % \ast

@@

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{lrlll}
      \textrm{Operators:}         & op   & {}\in Ops & {}::={} & + \mid * \\
      \textrm{Patterns:}          & q    & {}\in Pat & {}::={} & x \mid \hole \mid \{q|\ldots|q\} \\
      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda q \mathord{:}\tau.e \mid e~e \mid e~op~e \mid n \\
                                  &      &           &         & \phantom{x} \mid \hole \mid \{e|\ldots|e\} \\
      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \hole \mid \{ \tau | \ldots | \tau \} \\
    \end{array}
  \]
  \caption{Syntax as a grammar for the lambda calculus that we are considering.}
  \label{fig:Syntax as a grammar}
\end{figure}

@@

\[
  \begin{array}{l}
    \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
    \s : \G \to Act \times \V^{*} \times \V^{*} \times \V^{*}                                                                                          \\
    \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
    \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
    \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
    \CC(G) = \ldots                                                                                                                                    \\
    \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
    \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
    \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
    \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
    \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
    \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
  \end{array}
\]

@@

A \emph{graph} $G = [\e :-> s]_{u}^{*} \in \G = \A^{*}$ is a function from edges to edge states. Denote by $G(\e) = s$ the state $s$ of edge $\e$ in graph $G$, and by $G A$ the extension of graph $G$ by graph action $A$.

@@

A \emph{communication action} $a = (\Send, A^{*}, u^{*}) \in \A^{comm} = \{ \Send \} \times \A^{*} \times \U^{*}$ applies the graph action sequence $A^{*}$ to the editors identified by UUIDs $u^{*}$.

@@

%An \emph{editor} $E = (G,c,A_{Q}^{*},A_{H}^{*},u) \in \EE = \mathcal{G} \times \C \times \AA^{*} \times \AA^{*} \times \U$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{*}$, and known graph action history $A_{H}^{*}$, that can be identified by UUID $u$.

@@

An \emph{environment} $(E^{*}, A^{*}) \in \EE^{*} \times \AA^{*}$ is a set of communicating editors $E^{*}$ that share a global graph action history $A^{*}$.

@@

$\known(E_1 \cdots E_n) = A_{H1}^{*} \cdots A_{Hn}^{*}
  \iff (\forall j=1,\ldots,n)(E_j = (G_j,c_j,A_{Qj}^{*},A_{Hj}^{*},u_j))$

@@

\begin{mathpar}
  \inferrule{
  \leftIndex(i) = i'
  }{
  (G,(v,i),A_Q^{*},A_H^{*},u),A \xrightarrow{\Left} (G,(v,i'),A_Q^{*},A_H^{*},u),A
  }

  \inferrule{
  \rightIndex(i) = i'
  }{
  (G,(v,i),A_Q^{*},A_H^{*},u),A \xrightarrow{\Right} (G,(v,i'),A_Q^{*},A_H^{*},u),A
  }

  \inferrule{
  \parents(v) = \{(c',v',u')\}
  }{
  (G,(v,i),A_Q^{*},A_H^{*},u),A \xrightarrow{\Up} (G,c',A_Q^{*},A_H^{*},u),A
  }

  \inferrule{
  \cursorChildren(c) = \{(c',(k,u_k),u')\} \\
  \downIndex(k) = i
  }{
  (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Down} (G,((k,u_k),i),A_Q^{*},A_H^{*},u),A
  }

  \inferrule{}{
  (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Select~c'} (G,c',A_Q^{*},A_H^{*},u),A
  }
\end{mathpar}

@@

\begin{mathpar}
  \inferrule{
  \defaultIndex(c) = TODO(Was None) \\
  u, u_k \in \U \text{ fresh} \\
  A_H' = [(c,(k,u_k),u) :-> +]
  }{
  (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Create~k} (G A_H',c,A_Q^{*},A_H^{*}A_H',u),A
  }

  \inferrule{
  \defaultIndex(c) = i' \\
  c' = ((k,u_k),i') \\
  u, u_k \in \U \text{ fresh} \\
  \\\\
  \cursorChildren(G,c) = \{(c_j,v_j,u_j')\}_{j=1}^n \\
  A_H'^{*} =
    [(c  ,(k,u_k),u  ) :-> +]
    [(c' ,v_j    ,u_j') :-> +]_{j=1}^n
  [(c_j,v_j    ,u_j') :-> -]_{j=1}^n
  }{
  (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Create~k}
  (G A_H'^{*},c',A_Q^{*},A_H^{*}A_H'^{*},u),A
  }

  \inferrule{
  \cursorChildren(G,c) = \{\e_j\}_{j=1}^n \\
  A_H'^{*} = [\e_j :-> -]_{j=1}^n
  }{
  (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Destroy}
  (G A_H'^{*},c,A_Q^{*},A_H^{*}A_H'^{*},u),A
  }

  \inferrule{
  A_H = [(c,v,u') :-> +] \\
  u' \text{ fresh}
  }{
  (G,c,A_Q^{*},A_H^{*},u),A \xrightarrow{\Restore~v} (G A_H,c,A_Q^{*},A_H^{*}A_H,u),A %
  }
\end{mathpar}

@@

\paragraph{Record} \par
%
\begin{mathpar}
  \inferrule{
  a \in \A^{edit} \\
  (G,c,A_Q^{*},A_H^{*},u),A -->a (G',c',A_Q^{*},A_H'^{*},u),A
  }{
  (G,c,A_Q^{*},A_H^{*},u),A -->a (G',c',A_Q^{*}a,A_H'^{*},u),A
  }
\end{mathpar}

@@

$\boxed{E,A \xrightarrow{a^{*}} E,A}$
%
\begin{mathpar}
  \inferrule{
    E,A -->a E'',A'' \\
    E'',A'' \xrightarrow{a^{*}} E',A' \\
  }{
    E,A \xrightarrow{aa^{*}} E',A'
  }
\end{mathpar}

@@

$\boxed{(E,A)^{*} -->a (E,A)^{*}}$
%
\begin{mathpar}
  \inferrule{
  E_j = (G_j,c_j,A_{Qj}^{*},A_{Hj}^{*},u_j) \\
  E_j,A_j \xrightarrow{a'^{*}} (G_j',c_j',A_{Qj}'^{*}, A_{Hj}'^{*},u_j),A_j' \\
  E_j' =
  ( G_j'
  , c_j'
  , A_{Qj}'^{*} \setminus a'^{*}
  , A_{Hj}'^{*} \setminus \known(E_1 \cdots E_n),u_j) \\
  j = 1,\ldots,n
  }{
  (E_1,A_1) \cdots (E_n,A_n) \xrightarrow{\Send~ a'^{*} } (E_1',A_1') \cdots (E_n',A_n')
  }
\end{mathpar}

>>>>

% Load MnSymbol without clobbering \ast
% See https://tex.stackexchange.com/a/269691
\usepackage{amsmath}% needed before mathabx
\let\amsast=\ast
\usepackage[matha]{mathabx}% needed to prevent \ast getting clobbered
\let\abxast=\ast
\usepackage{MnSymbol}
\let\mnast=\ast
\let\ast=\abxast

@@

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{lrlll}
      \textrm{Operators:}         & op   & {}\in Ops & {}::={} & + \mid \ast{} \\
      \textrm{Patterns:}          & q    & {}\in Pat & {}::={} & x \mid \hole \mid \{q|\ldots|q\} \\
      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda q \mathord{:}\tau.e \mid e~e \mid e~op~e \mid n \\
                                  &      &           &         & \phantom{x} \mid \hole \mid \{e|\ldots|e\} \\
      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \hole \mid \{ \tau | \ldots | \tau \} \\
    \end{array}
  \]
  \caption{Syntax as a grammar for the lambda calculus that we are considering.}
  \label{fig:Syntax as a grammar}
\end{figure}

@@

\[
  \begin{array}{l}
    \textrm{User Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}}                                                       \\
    \s : \G \to Act \times \V^{\ast{}} \times \V^{\ast{}} \times \V^{\ast{}}                                                                                          \\
    \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G))                                                                                                         \\
    \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right.                                                           \\
    \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\}                                                                                   \\
    \CC(G) = \ldots                                                                                                                                    \\
    \D(G) = \orphans(G) \setminus \left\{ v_0 \right\}                                                                                                 \\
    \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right.                                                  \\
    \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\}                                                                                       \\
    \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
    \edges(G) = \left\{ e \mid G(e) \ne \bot \right\}                                                                                                  \\
    \liveEdges(G) = \left\{ e \mid G(e) = + \right\}                                                                                                   \\
  \end{array}
\]

@@

A \emph{graph} $G = [\e :-> s]_{u}^{\ast{}} \in \G = \A^{\ast{}}$ is a function from edges to edge states. Denote by $G(\e) = s$ the state $s$ of edge $\e$ in graph $G$, and by $G A$ the extension of graph $G$ by graph action $A$.

@@

A \emph{communication action} $a = (\Send, A^{\ast{}}, u^{\ast{}}) \in \A^{comm} = \{ \Send \} \times \A^{\ast{}} \times \U^{\ast{}}$ applies the graph action sequence $A^{\ast{}}$ to the editors identified by UUIDs $u^{\ast{}}$.

@@

%An \emph{editor} $E = (G,c,A_{Q}^{\ast{}},A_{H}^{\ast{}},u) \in \EE = \mathcal{G} \times \C \times \AA^{\ast{}} \times \AA^{\ast{}} \times \U$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{\ast{}}$, and known graph action history $A_{H}^{\ast{}}$, that can be identified by UUID $u$.

@@

An \emph{environment} $(E^{\ast{}}, A^{\ast{}}) \in \EE^{\ast{}} \times \AA^{\ast{}}$ is a set of communicating editors $E^{\ast{}}$ that share a global graph action history $A^{\ast{}}$.

@@

$\known(E_1 \cdots E_n) = A_{H1}^{\ast{}} \cdots A_{Hn}^{\ast{}}
  \iff (\forall j=1,\ldots,n)(E_j = (G_j,c_j,A_{Qj}^{\ast{}},A_{Hj}^{\ast{}},u_j))$

@@

\begin{mathpar}
  \inferrule{
  \leftIndex(i) = i'
  }{
  (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Left} (G,(v,i'),A_Q^{\ast{}},A_H^{\ast{}},u),A
  }

  \inferrule{
  \rightIndex(i) = i'
  }{
  (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Right} (G,(v,i'),A_Q^{\ast{}},A_H^{\ast{}},u),A
  }

  \inferrule{
  \parents(v) = \{(c',v',u')\}
  }{
  (G,(v,i),A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Up} (G,c',A_Q^{\ast{}},A_H^{\ast{}},u),A
  }

  \inferrule{
  \cursorChildren(c) = \{(c',(k,u_k),u')\} \\
  \downIndex(k) = i
  }{
  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Down} (G,((k,u_k),i),A_Q^{\ast{}},A_H^{\ast{}},u),A
  }

  \inferrule{}{
  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Select~c'} (G,c',A_Q^{\ast{}},A_H^{\ast{}},u),A
  }
\end{mathpar}

@@

\begin{mathpar}
  \inferrule{
  \defaultIndex(c) = TODO(Was None) \\
  u, u_k \in \U \text{ fresh} \\
  A_H' = [(c,(k,u_k),u) :-> +]
  }{
  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Create~k} (G A_H',c,A_Q^{\ast{}},A_H^{\ast{}}A_H',u),A
  }

  \inferrule{
  \defaultIndex(c) = i' \\
  c' = ((k,u_k),i') \\
  u, u_k \in \U \text{ fresh} \\
  \\\\
  \cursorChildren(G,c) = \{(c_j,v_j,u_j')\}_{j=1}^n \\
  A_H'^{\ast{}} =
    [(c  ,(k,u_k),u  ) :-> +]
    [(c' ,v_j    ,u_j') :-> +]_{j=1}^n
  [(c_j,v_j    ,u_j') :-> -]_{j=1}^n
  }{
  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Create~k}
  (G A_H'^{\ast{}},c',A_Q^{\ast{}},A_H^{\ast{}}A_H'^{\ast{}},u),A
  }

  \inferrule{
  \cursorChildren(G,c) = \{\e_j\}_{j=1}^n \\
  A_H'^{\ast{}} = [\e_j :-> -]_{j=1}^n
  }{
  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Destroy}
  (G A_H'^{\ast{}},c,A_Q^{\ast{}},A_H^{\ast{}}A_H'^{\ast{}},u),A
  }

  \inferrule{
  A_H = [(c,v,u') :-> +] \\
  u' \text{ fresh}
  }{
  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A \xrightarrow{\Restore~v} (G A_H,c,A_Q^{\ast{}},A_H^{\ast{}}A_H,u),A %
  }
\end{mathpar}

@@

\paragraph{Record} \par
%
\begin{mathpar}
  \inferrule{
  a \in \A^{edit} \\
  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A -->a (G',c',A_Q^{\ast{}},A_H'^{\ast{}},u),A
  }{
  (G,c,A_Q^{\ast{}},A_H^{\ast{}},u),A -->a (G',c',A_Q^{\ast{}}a,A_H'^{\ast{}},u),A
  }
\end{mathpar}

@@

$\boxed{E,A \xrightarrow{a^{\ast{}}} E,A}$
%
\begin{mathpar}
  \inferrule{
    E,A -->a E'',A'' \\
    E'',A'' \xrightarrow{a^{\ast{}}} E',A' \\
  }{
    E,A \xrightarrow{aa^{\ast{}}} E',A'
  }
\end{mathpar}

@@

$\boxed{(E,A)^{\ast{}} -->a (E,A)^{\ast{}}}$
%
\begin{mathpar}
  \inferrule{
  E_j = (G_j,c_j,A_{Qj}^{\ast{}},A_{Hj}^{\ast{}},u_j) \\
  E_j,A_j \xrightarrow{a'^{\ast{}}} (G_j',c_j',A_{Qj}'^{\ast{}}, A_{Hj}'^{\ast{}},u_j),A_j' \\
  E_j' =
  ( G_j'
  , c_j'
  , A_{Qj}'^{\ast{}} \setminus a'^{\ast{}}
  , A_{Hj}'^{\ast{}} \setminus \known(E_1 \cdots E_n),u_j) \\
  j = 1,\ldots,n
  }{
  (E_1,A_1) \cdots (E_n,A_n) \xrightarrow{\Send~ a'^{\ast{}} } (E_1',A_1') \cdots (E_n',A_n')
  }
\end{mathpar}
